---
title: "Data607_Project3_Analysis"
author: " John Ferrara, Alinzon Simon, Akeem Lawrence, Anthony Roman, and Ben Wolin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RMySQL)
library(readr)
library(tidyr)
library(dplyr)
```



## Connecting to the Local Sql DB That was created via the Loading Script (./loading/job_post.sql)

```{r sqlConnection}

# Read the password in from file
# ---- This will need to be changed for others if we dont use cloud infrastructure. 
pw_path <- "~/Documents/secrets/mysql_root_pw.txt"
password <- read_file(path.expand(pw_path))

con <- dbConnect(RMySQL::MySQL(),
                 dbname = "job_portal",
                 host = "127.0.0.1",
                 port = 3306,
                 user = "root",
                 password = password,
                client.flag = CLIENT_LOCAL_FILES)

```

## Loading the Data (./data) into working memory here for processing.

You can also embed plots, for example:

```{r data_loading}

#Reading in job_postings (./data/raw/job_postings.csv)
job_postings_path <- "./data/raw/job_postings.csv"
job_postings <- read.csv(job_postings_path)

#Checking it looks ok 
print(head(job_postings))
print(colnames(job_postings))
 # [1] "job_link"            "last_processed_time" "last_status"         "got_summary"         "got_ner"            
 # [6] "is_being_worked"     "job_title"           "company"             "job_location"        "first_seen"         
# [11] "search_city"         "search_country"      "search_position"     "job_level"           "job_type" 

#Reading in job_skills (./data/raw/job_skills.csv)
job_skills_path <- "./data/raw/job_skills.csv"
job_skills <- read.csv(job_skills_path)
#Checking it looks ok 
print(head(job_skills))
print(colnames(job_skills))
# [1] "job_link"   "job_skills"

#Reading in job_summary (./data/raw/job_summary.csv)
job_summary_path <- "./data/raw/job_summary.csv"
job_summary <- read.csv(job_summary_path)
#Checking it looks ok
print(head(job_summary))
print(colnames(job_summary))
# [1] "job_link"    "job_summary"

```
## Parsing and Processing Job Posting Data into tables to match Entity-Entity Relationship DIagram (./additonal_materials/EER.mwb) 

### Working with Job Postings
```{r job_postings}


# Looking at the job status data in the job_postings data 
print(unique(job_postings$last_status))
## not ideal there is only one value in status; Making table anyway

## Creating Status Table Df 
status_values <- unique(job_postings$last_status)
id_vector <-seq(from = 1, to = length(status_values), by =1) 
status_table_df <- data.frame("status_id" = id_vector,
                              "status_name"=status_values)
status_table_df
write.csv(status_table_df, "data/processed/status_table.csv")

## not ideal there is only one value in status; 

### Haveing Issues loading to local MySQL, also not sure wht platform we are using in cloud so just putting data in csvs atm.
# dbWriteTable(con, name = "status", value = status_table_df, append = TRUE, row.names = FALSE)


## Creating Location Table Df
### NOTE: there is no state column, and nothing to parse out of the city column as the Kaggle Source page suggests.
print(unique(job_postings$search_city))    
print(unique(job_postings$search_country)

location_df <- job_postings %>%
    select(search_city, search_country)
location_df <- distinct(location_df)
#Adding Null State COlumn to match SQL table 
location_df[ , 'state'] = NA
location_df <- location_df %>% 
  rename(country = search_country,city = search_city)
#Adding id vector for final df 
location_df$location_id <- seq(from = 1, to = nrow(location_df), by =1) 
location_table_df<- location_df[,c("location_id","city","state","country")]
write.csv(location_table_df, "data/processed/location_table.csv")

##Creating Company Table DF 
# print(unique(job_postings$company))
company_values <- unique(job_postings$company)
id_vector <-seq(from = 1, to = length(company_values), by =1) 
company_table_df <- data.frame("company_id" = id_vector,
                               "company_name"=company_values)
write.csv(company_table_df, "data/processed/company_table.csv")

## Job Post Id interim df
link_values <- unique(job_postings$job_link)
id_vector <-seq(from = 1, to = length(link_values), by =1) 
link_id_table_df <- data.frame("job_post_id" = id_vector,
                               "job_post_link"=link_values)


### Adding Job Summary to the job posting df
job_postings_w_summary <- job_postings %>% left_join(job_summary, by=c('job_link'))
## Adding COmpany ID
temp_comp <-company_table_df %>% rename(company=company_name)
job_postings_w_summary <- job_postings_w_summary %>% left_join(temp_comp, by=c('company'))
## Adding Status id
status_comp <- status_table_df %>% rename(last_status=status_name)
job_postings_w_status <- job_postings_w_summary %>% left_join(status_comp, by=c('last_status'))
## adding location id
loc_temp <- location_df[,c("location_id","city","country")]
loc_temp <- loc_temp%>%rename(search_city=city,search_country=country)
job_postings_w_loc <- job_postings_w_status %>% left_join(loc_temp, by=c('search_city','search_country'))
### Adding Job post id (table used made down below made first)
job_post_id_temp <- skill_df_link_flat_id[,c("job_post_id","job_post_link")]
job_post_id_temp<- job_post_id_temp%>%rename(job_link = job_post_link)
job_posting_w_all <- job_postings_w_loc %>% left_join(job_post_id_temp, by=c('job_link'))

## Limiting and renaming to make the job_post table
job_posting_table_df <- job_posting_w_all[,c("job_post_id","company_id","status_id","location_id","job_title","job_summary","last_processed_time")]
write.csv(job_posting_table_df, "data/processed/job_posting_table.csv")


```

### Working with Job Skill File
``` {r job_skills}
### This df will be needed to add ids to the job post skill table via links
skill_df_link_flat <- data.frame(job_link = character(), skill = character(), stringsAsFactors = FALSE)
for (i in 1:nrow(job_skills)) {
  # print(i)
  ink_val <- job_skills[i, "job_link"]
  skills_vec <-strsplit(job_skills[i, "job_skills"],",")[[1]]
  for (skill in skills_vec) {
    # strtrim(skill)
    skill_df_link_flat <- rbind(skill_df_link_flat, data.frame(job_link = link_val, skill = trimws(skill), stringsAsFactors = FALSE))
  }
}

#Creating SKill Table 
skill_values <- unique(skill_df_link_flat$skill)
id_vector <-seq(from = 1, to = length(skill_values), by =1) 
skill_table_df <- data.frame("skill_id" = id_vector,
                               "skill_name"=skill_values)
write.csv(skill_table_df, "data/processed/skill_table.csv")


## Left join skill id to the skill_df_link_flat
skill_df_link_flat<- skill_df_link_flat %>% rename(skill_name = skill,job_post_link=job_link)
skill_df_link_flat_id <- skill_df_link_flat %>% left_join(skill_table_df, by=c('skill_name'))
## Joining in the job post id
skill_df_link_flat_id <- skill_df_link_flat_id %>% left_join(link_id_table_df, by=c('job_post_link'))

### Creating job_post_skill table df
job_post_skill_temp <-distinct(skill_df_link_flat_id[,c("job_post_id","skill_id")])
id_vector <-seq(from = 1, to = nrow(job_post_skill_temp), by =1) 
job_post_skill_df <- data.frame("id" = id_vector,
                                "job_post_id"=job_post_skill_temp$job_post_id,
                               "skill_id"=job_post_skill_temp$skill_id)
write.csv(job_post_skill_df, "data/processed/job_post_skill.csv")

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
